### 2025-02-16 src/paper/paper.tex
minor changes

Table~\ref{tab:preprocessing_pipeline} presents the preprocessing techniques used and their respective hyperparameters and tools used. Figure~\ref{fig:slice_comparison} gives a comparison of a raw image and a preprocessed image, also indicating the image metrics.
%Figure~\ref{fig:test_pipeline} gives a visual representation of the conducted tests.
%
\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\textwidth]{./figs/slice_comparison.png} % test_pipeline.png
    \caption{Panel A shows a slice of a raw scan. Panel B shows a slice of a preprocessed scan. For preprocessed scans, the standard deviation of SNR values is relatively low (\mbox{$SNR= 25 \pm 1$}), which ensures consistency in the image quality and prevents the model from making predictions based on possible artifacts.}\label{fig:slice_comparison}
\end{figure}

@@

The results presented in Table~\ref{tab:model_meetrics_augmentation} highlight the impact of various data augmentation methods on model performance. The baseline model using raw data (0) achieved an accuracy of 0.51 and an F1-score of 0.63. In contrast, applying translation (6) significantly improved both accuracy (75\%) and F1-score (0.79), with a notable increase in recall (0.84), indicating that translation is effective at boosting model performance and sensitivity.

Rotation (7) further improved results, with the highest accuracy (79\%) and F1-score (0.80), alongside strong recall (0.81) and precision (0.78), demonstrating its ability to enhance both model robustness and overall performance.

Gaussian noise (8), however, led to a substantial decrease in performance, with an accuracy of only 33\% and a low F1-score of 0.40. This suggests that adding noise may negatively affect the model, potentially obscuring relevant patterns in the data.
\begin{center}
	\begin{table}[t]
		\centering
		\caption{\label{tab:model_meetrics_augmentation}Classification performance metrics for different data augmentation methods. Accuracy, recall, precision, and F1-score are reported for models trained on raw MRI slices and augmented datasets. Both translation (6) and rotation (7) significantly improved model performance, with rotation achieving the highest accuracy (79\%) and F1-score (0.80). In contrast, Gaussian noise (8) led to a severe drop in performance, reducing accuracy to 33\% and F1-score to 0.40. The final training setup, combining preprocessing with translation and rotation (1+2+3+6+7), achieved the best overall results, with 81\% accuracy and the highest recall (0.98), indicating the effectiveness of spatial transformations in improving classification.}
		\begin{tabular*}{500pt}{@{\extracolsep\fill}lcccc@{\extracolsep\fill}}
			\toprule
			\textbf{Augmentation method} &\textbf{Accuracy (\%)} &  \textbf{Recall} &\textbf{Precision} &\textbf{F1-score}\\
			\midrule
			Raw (0) & 0.51 & 0.67 & 0.60 & 0.63 \\
			Translation (6) & 75\% & 0.84 & 0.74 & 0.79 \\
			Rotation (7) & 79\% & 0.81 & 0.78 & 0.80 \\
			Gaussian Noise (8) & 33\% & 0.32 & 0.53 & 0.40 \\
            Final training (1+2+3+6+7) & 81\% & 0.98 & 0.77 & 0.86 \\
			\bottomrule
		\end{tabular*}
	\end{table}
\end{center}

The final training combined the optimal combination of preprocessing with translation and rotation techniques (1+2+3+6+7), surpassing the accuracy and recall of all the experiments (81\% and 0.98 respectively), with a precision value only surpassed by the rotation experiment. Figure~\ref{fig:eval_final_model} shows the output of the final training and Figure~\ref{fig:model_metrics_comparison} shows the comparison between the metrics of all the preprocessing and augmentation experiments.
\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{./figs/eval_final_model.png} % 
    \caption{Preprocessing and augmentation steps: resampling + normalization + brain extraction + translation + rotation. Panel A: confusion matrix with recall of 0.98. Panel B: ROC curve with AUC of 0.86}\label{fig:eval_final_model}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{./figs/model_metrics_comparison.png} % 
    \caption{Summary of the metrics for each preprocessing experiment. Panel A: comparison of accuracies across the experiments. The highest accuracy was recorded in the final experiment combining resampling, normalization, brain extraction, translation and rotation. Panel B: comparison of recall scores. The final experiment yielded the highest recall score. Panel C: comparison of precision scores. The rotation experiment produced the highest score, probably due to an increase in instances of the negative class after augmentation. The numbers along the y-axis represent preprocessing and augmentation scenarios: 0 --- raw images, 1 --- resampling, 2 --- normalization, 3 --- brain extraction, 4 --- cropping, 5 --- smoothing, 6 --- translation, 7 --- rotation, 8 --- Gaussian noise.}\label{fig:model_metrics_comparison}
\end{figure}

>>>>

Table~\ref{tab:preprocessing_pipeline} presents the preprocessing techniques used and their respective hyperparameters and tools used. Figure~\ref{fig:slice_comparison} gives a comparison of a raw image and a preprocessed image, also indicating the image metrics.
%Figure~\ref{fig:test_pipeline} gives a visual representation of the conducted tests.
%
\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\textwidth]{./figs/slice_comparison.png} % test_pipeline.png
    \caption{Panel A shows a slice of a raw scan. Panel B shows a slice of a preprocessed scan. For preprocessed scans, the standard deviation of SNR values is relatively low (\mbox{$SNR= 25 \pm 2$}), which ensures consistency in the image quality and prevents the model from making predictions based on possible artifacts.}\label{fig:slice_comparison}
\end{figure}

@@

The results presented in Table~\ref{tab:model_meetrics_augmentation} highlight the impact of various data augmentation methods on model performance. The baseline model using raw data (0) achieved an accuracy of 0.51 and an F1-score of 0.63. In contrast, applying translation (6) significantly improved both accuracy (75\%) and F1-score (0.79), with a notable increase in recall (0.84), indicating that translation is effective at boosting model performance and sensitivity.
\begin{center}
	\begin{table}[t]
		\centering
		\caption{\label{tab:model_meetrics_augmentation}Classification performance metrics for different data augmentation methods. Accuracy, recall, precision, and F1-score are reported for models trained on raw MRI slices and augmented datasets. Both translation (6) and rotation (7) significantly improved model performance, with rotation achieving the highest accuracy (79\%) and F1-score (0.80). In contrast, Gaussian noise (8) led to a severe drop in performance, reducing accuracy to 33\% and F1-score to 0.40. The final training setup, combining preprocessing with translation and rotation (1+2+3+6+7), achieved the best overall results, with 81\% accuracy and the highest recall (0.98), indicating the effectiveness of spatial transformations in improving classification.}
		\begin{tabular*}{500pt}{@{\extracolsep\fill}lcccc@{\extracolsep\fill}}
			\toprule
			\textbf{Augmentation method} &\textbf{Accuracy (\%)} &  \textbf{Recall} &\textbf{Precision} &\textbf{F1-score}\\
			\midrule
			Raw (0) & 0.51 & 0.67 & 0.60 & 0.63 \\
			Translation (6) & 75\% & 0.84 & 0.74 & 0.79 \\
			Rotation (7) & 79\% & 0.81 & 0.78 & 0.80 \\
			Gaussian Noise (8) & 33\% & 0.32 & 0.53 & 0.40 \\
            Final training (1+2+3+6+7) & 81\% & 0.98 & 0.77 & 0.86 \\
			\bottomrule
		\end{tabular*}
	\end{table}
\end{center}
Rotation (7) further improved results, with the highest accuracy (79\%) and F1-score (0.80), alongside strong recall (0.81) and precision (0.78), demonstrating its ability to enhance both model robustness and overall performance.

Gaussian noise (8), however, led to a substantial decrease in performance, with an accuracy of only 33\% and a low F1-score of 0.40. This suggests that adding noise may negatively affect the model, potentially obscuring relevant patterns in the data.
\begin{figure}
    \centering
    \includegraphics[width=1.0\textwidth]{./figs/eval_final_model.png} % 
    \caption{Model evaluation for the final training. Preprocessing and augmentation steps: resampling + normalization + brain extraction + translation + rotation. Panel A: confusion matrix with recall of 0.98. Panel B: ROC curve with AUC of 0.86}\label{fig:eval_final_model}
\end{figure}

The final training combined the optimal combination of preprocessing with translation and rotation techniques (1+2+3+6+7), surpassing the accuracy and recall of all the experiments (81\% and 0.98 respectively), with a precision value only surpassed by the rotation experiment. Figure~\ref{fig:eval_final_model} shows the output of the final training and Figure~\ref{fig:model_metrics_comparison} shows the comparison between the metrics of all the preprocessing and augmentation experiments.
\begin{figure}
    \centering
    \includegraphics[width=1.0\textwidth]{./figs/model_metrics_comparison.png} % 
    \caption{Summary of the metrics across different experiments. Panel A: comparison of accuracies across the experiments. The highest accuracy was recorded in the final experiment combining resampling, normalization, brain extraction, translation and rotation. Panel B: comparison of recall scores. The final experiment yielded the highest recall score. Panel C: comparison of precision scores. The rotation experiment produced the highest score, probably due to an increase in instances of the negative class after augmentation. The numbers along the y-axis represent preprocessing and augmentation scenarios: 0 --- raw images, 1 --- resampling, 2 --- normalization, 3 --- brain extraction, 4 --- cropping, 5 --- smoothing, 6 --- translation, 7 --- rotation, 8 --- Gaussian noise.}\label{fig:model_metrics_comparison}
\end{figure}
